---
title: "Scores-Selections Comparison"
author: "Michel Tarnow"
date: today
format: 
    html:
      toc: true
      self-contained: true
      highlight-style: github
      code-line-numbers: true
editor_options: 
    chunk_output_type: inline
---

## Import Packages

```{r}
suppressPackageStartupMessages({
  library(reshape2)
  library(MLmetrics)
  library(caret)
  library(ComplexUpset)
  library(patchwork)
  library(scran)
  library(scater)
  library(igraph)
  library(SingleCellExperiment)
  library(zellkonverter)
  library(MLVSBM)
  library(tidyverse)
  library(viridis)
  library(GGally)
  library(rlang)
  library(colorspace)
})
```



## Scores

### Helper Functions

```{r}
# helper function for loading and processing Patches scores
load_process_patches_scores <- function(file) {
  
  # read in df from file and fix column names
  df <- read_csv(
    file,
    col_names = TRUE,
    name_repair = "minimal",
    show_col_types = FALSE
  )
  colnames(df)[1] = "gene_id"
  
  # remove _Patches from column names and
  # replace spaces in column names with underscores and
  # remove + signs
  df <- df %>%
    rename_with(~ sub("_Patches$", "", .x), ends_with("_Patches")) %>% 
    rename_with(~ str_replace_all(.x, " ", "_")) %>% 
    rename_with(~ str_replace_all(.x, "\\+", ""))
  
  # add combinations of scores
  
  ## max absolute value for condition (treatment)
  df <- df %>% 
    mutate(
      sMAV = pmax(
        abs(stim_score),
        abs(ctrl_score)
      )
    )
  
  ## Euclidean norm for condition (treatment)
  df <- df %>% 
    mutate(
      sEN = sqrt(
        stim_score^2 +
          ctrl_score^2
      )
    )
  
  ## absolute difference for condition
  df <- df %>% 
    mutate(
      sAD = abs(stim_score - ctrl_score))
    
  if ("B_cells_score" %in% names(df)) {
    # cell type and condition scores available
    
    exclude_cols <- c("common_score", "stim_score", "ctrl_score")
    
    ## max absolute value for cell types
    df <- df %>% 
      mutate(
        tMAV = pmax(
          abs(B_cells_score),
          abs(CD14_Monocytes_score),
          abs(CD4_T_cells_score),
          abs(CD8_T_cells_score),
          abs(Dendritic_cells_score),
          abs(FCGR3A_Monocytes_score),
          abs(Megakaryocytes_score),
          abs(NK_cells_score)
        )
      )
    
    ## Euclidean norm for cell types
    df <- df %>% 
      mutate(
        tEN = sqrt(
          B_cells_score^2 +
            CD14_Monocytes_score^2 +
            CD4_T_cells_score^2 +
            CD8_T_cells_score^2 +
            Dendritic_cells_score^2 +
            FCGR3A_Monocytes_score^2 +
            Megakaryocytes_score^2 +
            NK_cells_score^2
        )
      )
    
    ## mean absolute difference for group
    df <- df %>% 
      mutate(
        tMAD = {
          # select all relevant columns
          score_cols <- grep("cells|cytes", names(df), value = TRUE)
          
          # get all unique pairs of columns
          pairs <- combn(score_cols, 2, simplify = FALSE)
          
          # compute all absolute pairwise differences
          diffs <- lapply(pairs, function(p) abs(df[[p[1]]] - df[[p[2]]]))
          
          # take rowMeans of those differences
          rowMeans(do.call(cbind, diffs), na.rm = TRUE)
        }
      )
    
  }
  
  return(df)
    
}

# helper function for loading and processing FSS scores
load_process_fss_scores <- function(files) {
  
  # select gene and score columns and combine all dfs in one list
  dfs <- lapply(files, function(file) {
    
    # read in df from file
    df <- readRDS(file)
    
    # rename score value column to score name
    if (nrow(df) > length(unique(df$gene_id))) {
      # df contains both tPVE and sPVE scores
      
      df <- df %>% 
        pivot_wider(names_from = sco, values_from = sco_val)
      cols_to_keep <- c("gene_id", "tPVE", "sPVE")
      
    } else {
      # df contains a single score (all but tPVE + sPVE)
      
      new_colname <- df$sco[1]
      df <- df %>% 
        rename(!!new_colname := sco_val)
      cols_to_keep <- c("gene_id", new_colname)
      
    }
    
    # select gene and score columns
    df <- df %>% 
      select(all_of(cols_to_keep))
    
    # return df
    return(df)
    
  })
  
  # combine all dfs in one df
  df_combined <- as_tibble(reduce(
    dfs, 
    ~ inner_join(.x, .y, by = "gene_id")
  ))
  
  # return combined df
  return(df_combined)
  
}
```

### Load and Process Patches Scores

```{r}
files <- list.files(
  path = "../../data/kang/02-scores_patches/",
  pattern = "\\.csv$", full.names = TRUE,
)

patches_scores <- lapply(files, load_process_patches_scores)
names(patches_scores) <- sub("_loadings.csv", "", basename(files))
```

```{r}
lapply(patches_scores, head)
```

### Load and Process FSS Scores

```{r}
files <- list.files(
  "../../data/kang/02-scores_fss/",
  pattern = "\\.rds$", 
  full.names = TRUE
)

fss_scores <- load_process_fss_scores(files)
```

```{r}
head(fss_scores)
```

### Combine and Transform Scores

```{r}
# combine Patches and FSS scores in one df for each Patches workflow
# and add z-transformed scores
# and add ranks
all_scores <- lapply(patches_scores, function(df) {
  
  new_df <- cbind.data.frame(
    df,
    fss_scores %>% 
      select(!gene_id)
  )
  
  # add z-transformed scores and ranks
  if ("common_score_z" %in% colnames(df)) {
    # z-transformations already exist, skip df
    return(new_df)
  }
  
  new_df <- new_df %>% 
    mutate(
      across(
        .cols = -gene_id,
        .fns = list(
          z = ~ as.numeric(scale(.x)),
          rank = ~ rank(-.x, ties.method = "first")
        ),
        .names = "{.col}_{.fn}"
      )
    )
  
  return(new_df)
  
})

lapply(all_scores, head)
```

### Compare Scores

#### Correlation Between Scores

```{r}
# create correlation plots
cor_plots_rank <- lapply(all_scores, function(df) {
  
  # select z-transformed scores
  df_rank <- df %>% 
    select(ends_with("_rank"))
  
  # calculate pairwise Spearman rank correlations
  cor_mat <- cor(
    df_rank,
    use = "pairwise.complete.obs",
    method = "spearman"
  )
  
  # plot correlations as a heat map
  cor_long <- melt(cor_mat)
  ggplot(data = cor_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(-1, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Correlation")
  
})

cor_plots_rank
```

```{r}
# combine individual plots
subtitles <- c(
  "Cell Types & Conditions",
  "Conditions"
)
cor_plots_rank <- Map(function(p, sub) p + ggtitle(sub), cor_plots_rank, subtitles)

cor_plots_rank <- lapply(cor_plots_rank, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(cor_plots_rank, ncol = 1, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Ranked Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_rank.png", combined_plot, path = "../../outs/kang/",
       width = 10, height = 14, dpi = 300)
combined_plot
```

```{r}
# create correlation plots
cor_plots_z <- lapply(all_scores, function(df) {
  
  # select z-transformed scores
  df_z <- df %>% 
    select(ends_with("_z"))
  
  # calculate pairwise Spearman rank correlations
  cor_mat <- cor(
    df_z,
    use = "pairwise.complete.obs",
    method = "spearman"
  )
  
  # plot correlations as a heat map
  cor_long <- melt(cor_mat)
  ggplot(data = cor_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(-1, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Correlation")
  
})

cor_plots_z
```

```{r}
# combine individual plots
subtitles <- c(
  "Cell Types & Conditions",
  "Conditions"
)
cor_plots_z <- Map(function(p, sub) p + ggtitle(sub), cor_plots_z, subtitles)

cor_plots_z <- lapply(cor_plots_z, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(cor_plots_z, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Z-Transformed Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_z.png", combined_plot, path = "../../outs/kang/",
       width = 18, height = 8, dpi = 300)
combined_plot
```

#### Overlap of top-n genes (Jaccard Index)

```{r}
# function to compute the top-n genes for a given rank
get_top_n_genes <- function(df, rank, n) {
  top_genes <- (df %>% 
    arrange(.data[[rank]]) %>% 
    slice_head(n = n))$gene_id
  
  return(top_genes)
}

# function to compute Jaccard index for top n genes
compute_jaccard <- function(df, rank1, rank2, n) {
  top1 <- get_top_n_genes(df, rank1, n)
  top2 <- get_top_n_genes(df, rank2, n)
  length(intersect(top1, top2)) / length(union(top1, top2))
}

# function to compute Jaccard matrix for all provided ranks
compute_jaccard_matrix <- function(df, n) {
  
  rank_cols <- grep("_rank$", colnames(df), value = TRUE)
  m <- length(rank_cols)
  
  jmat <- matrix(NA, nrow = m, ncol = m, dimnames = list(rank_cols, rank_cols))
  
  for (i in seq_along(rank_cols)) {
    for (j in seq_along(rank_cols)) {
      jmat[i, j] <- compute_jaccard(df, rank_cols[i], rank_cols[j], n)
    }
  }
  
  return(jmat)
  
}
```

```{r}
# plot Jaccard indices as heat maps
j_plots <- lapply(all_scores, function(df) {
  
  j_mat <- compute_jaccard_matrix(df, 100)
  
  j_long <- melt(j_mat)
  ggplot(data = j_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(0, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Jaccard\nIndex")
  
})

j_plots
```

```{r}
# combine individual plots
subtitles <- c(
  "Cell Types & Conditions",
  "Conditions"
)
j_plots <- Map(function(p, sub) p + ggtitle(sub), j_plots, subtitles)

j_plots <- lapply(j_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(j_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Jaccard Index of Top-100 Genes",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_jaccard.png", combined_plot, path = "../../outs/kang/",
       width = 18, height = 8, dpi = 300)
combined_plot
```



## Selections

### Load and Process FSS Selections

```{r}
load_process_fss_selections <- function(files) {
  
  # select gene id and selection columns and combine all dfs in one list
  fss_selection_dfs <- lapply(files, function(file) {
    
    # read in df from file
    df <- readRDS(file)
    
    # df contains a single selection
    new_colname <- df$sel[1]
    df <- df %>% 
      rename(!!new_colname := sel_val)
    cols_to_keep <- c("gene_id", new_colname)
    
    # select gene and score columns
    df <- df %>% 
      select(all_of(cols_to_keep))
    
    return(df)
    
  })
  
  # combine all dfs in one df
  fss_selections <- as_tibble(reduce(
    fss_selection_dfs, 
    ~ inner_join(.x, .y, by = "gene_id")
  ))
  
  # sort columns
  fss_selections <- fss_selections %>% 
    select(everything())
  
  return(fss_selections)
  
}
```

```{r}
# get all files containing selections
rds_files <- list.files(
  "../../data/kang/03-selections_fss/",
  pattern = "\\.rds$", 
  full.names = TRUE
)

# load and process FSS selections
fss_selections <- load_process_fss_selections(rds_files)

head(fss_selections)
```

```{r}
# number of genes to select
n <- sum(fss_selections$tF)
```

```{r}
# add new FSS selections
expand_fss_selections <- function(df) {
  
  # get FSS scores
  score_df <- fss_scores
  
  # tF and sPBDS
  score_names <- c("tF", "sPBDS")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    df$sPBDS_tF <- scores >= cutoff
    
  }
  
  # tF and sPVE
  score_names <- c("tF", "sPVE")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    df$sPVE_tF <- scores >= cutoff
    
  }
  
  # tPVE and sPVE
  score_names <- c("tPVE", "sPVE")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    df$sPVE_tPVE <- scores >= cutoff
    
  }
  
  return(df)
  
}

fss_selections <- expand_fss_selections(fss_selections)
head(fss_selections)
```

### Compute Patches Selections

```{r}
patches_selections <- lapply(patches_scores, function(score_df) {
  
  # create binary df that contains binary selections based on scores
  binary_df <- score_df
  for (col in names(score_df)) {
    
    # skip gene id column
    if (col == "gene_id") {
      next
    }
    
    # convert score into binary predictions
    scores <- score_df[[col]]
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df[[col]] <- scores >= cutoff
    
  }
  
  # add new selections based on combinations of scores
  
  # difference in rank of max absolute group and condition scores
  max_abs <- c("tMAV", "sMAV")
  if (all(max_abs %in% colnames(score_df))) {
    
    t <- rank(score_df[[max_abs[1]]])
    s <- rank(score_df[[max_abs[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$tMAV_sMAV <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sMAV_tMAV <- scores >= cutoff
    
  }
  
  # difference in rank of Euclidean norm group and condition scores
  eucl_norm <- c("tEN", "sEN")
  if (all(eucl_norm %in% colnames(score_df))) {
    
    t <- rank(score_df[[eucl_norm[1]]])
    s <- rank(score_df[[eucl_norm[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$tEN_sEN <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sEN_tEN <- scores >= cutoff
    
  }
  
  # difference in rank of absolute difference group and condition scores
  abs_dif <- c("tMAD", "sAD")
  if (all(abs_dif %in% colnames(score_df))) {
    
    t <- rank(score_df[[abs_dif[1]]])
    s <- rank(score_df[[abs_dif[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$tMAD_sAD <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sAD_tMAD <- scores >= cutoff
    
  }
  
  # difference in rank of common score and condition scores
  com_con <- c("common_score", "sMAV", "sEN", "sAD")
  if (all(com_con %in% colnames(score_df)) && !("B_cells_score" %in% colnames(score_df))) {
    
    t <- rank(score_df[[com_con[1]]])
    s1 <- rank(score_df[[com_con[2]]])
    s2 <- rank(score_df[[com_con[3]]])
    s3 <- rank(score_df[[com_con[4]]])
    
    # common - sMAV
    scores <- t - s1
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$common_sMAV <- scores >= cutoff
    
    # common - sEN
    scores <- t - s2
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$common_sEN <- scores >= cutoff
    
    # common - sAD
    scores <- t - s3
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$common_sAD <- scores >= cutoff
    
    # sMAV - common
    scores <- s1 - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sMAV_common <- scores >= cutoff
    
    # sEN - common
    scores <- s2 - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sEN_common <- scores >= cutoff
    
    # sAD - common
    scores <- s3 - t
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df$sAD_common <- scores >= cutoff
    
  }
  
  # remove selections not needed for downstream analyses
  if ("B_cells_score" %in% colnames(binary_df)) {
    binary_df <- binary_df %>% 
      select(
        -B_cells_score,
        -CD14_Monocytes_score,
        -CD4_T_cells_score,
        -CD8_T_cells_score,
        -Dendritic_cells_score,
        -FCGR3A_Monocytes_score,
        -Megakaryocytes_score,
        -NK_cells_score
      )
  }
  binary_df <- binary_df %>% 
    select(
      -ctrl_score,
      -stim_score
    )
  
  return(binary_df)
  
})

names(patches_selections) <- names(patches_scores)
lapply(patches_selections, head)
```

### Combine Selections

```{r}
# combine Patches and FSS selections in one df for each Patches workflow
all_selections <- lapply(patches_selections, function(df) {
    
  new_df <- cbind.data.frame(
    df,
    fss_selections %>% 
      select(!gene_id)
  )
  
  # add all selection as baseline for downstream analyses
  new_df$all_genes <- TRUE
  
  return(new_df)
  
})

lapply(all_selections, head)
```

### Compare Selections

#### Overlaps

```{r}
plot_overlaps <- function(df, title) {
  
  # FSS selections
  fss_cols <- c(
    "random",
    "HVG",
    "tF",
    "tF_sPBDS",
    "tF_sPVE",
    "tPVE_sPVE",
    "sPBDS_tF",
    "sPVE_tF",
    "sPVE_tPVE"
  )
  
  # Patches selections
  patches_cols <- setdiff(colnames(df), c(fss_cols, "all_genes"))

  # compute overlaps
  overlap_df <- expand.grid(Patches = patches_cols, FSS = fss_cols, stringsAsFactors = FALSE) %>% 
    rowwise() %>% 
    mutate(Overlap = sum(df[[Patches]] & df[[FSS]])) %>% 
    ungroup()
  
  # order scores for plot
  overlap_df$Patches <- factor(overlap_df$Patches, levels = patches_cols)
  
  # plot overlap as heat map
  ggplot(overlap_df, aes(x = Patches, y = FSS, fill = Overlap)) +
    geom_tile(color = "white") +
    geom_text(aes(label = Overlap), color = "black") +
    scale_fill_viridis_c(option = "C", direction = -1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, x = "", y = "")
  
}
```

```{r}
p1 <- plot_overlaps(
  all_selections$Kang18_clu_con %>% select(-gene_id),
  "Cell Types & Conditions"
)
p1
```

```{r}
p2 <- plot_overlaps(
  all_selections$Kang18_con %>% select(-gene_id),
  "Conditions"
)
p2
```

```{r}
# combine individual plots
overlap_plots <- list(p1, p2)
subtitles <- c(
  "Cell Types & Conditions",
  "Conditions"
)
overlap_plots <- Map(function(p, sub) p + ggtitle(sub), overlap_plots, subtitles)

overlap_plots <- lapply(overlap_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(overlap_plots, ncol = 1, guides = "collect") +
  plot_annotation(
    title = "Overlap of Selected Genes (n = 2000)",
    theme = theme(
      legend.position = "none",
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_overlap.png", combined_plot, path = "../../outs/kang/",
       width = 8, height = 9, dpi = 300)
combined_plot
```

### Export Selections

```{r}
# read in sce file
sce <- readRDS("../../data/kang/01-pro/Kang18.rds")

# function for adding selection to sce objects
add_selection_data <- function(sce, selections_list, name) {
  
  selection <- selections_list[[name]]
  if (!all(rownames(sce) == selection$gene_id)) {
    stop("Gene IDs do not match between SCE and selection table")
  }
  
  rowData(sce) <- cbind(rowData(sce), selection[ , -1])
  return(sce)
  
}

# add selections to sce objects
sce_clu_con <- add_selection_data(sce, all_selections, "Kang18_clu_con")
sce_con     <- add_selection_data(sce, all_selections, "Kang18_con")

# export sce to anndata
writeH5AD(sce_clu_con, "../../data/kang/03-selections/Kang18-selections_clu_con.h5ad")
writeH5AD(sce_con, "../../data/kang/03-selections/Kang18-selections_con.h5ad")
```



## Downstream Analyses

### Load SingleCellExperiment

```{r}
sce <- readRDS("../../data/kang/01-pro/Kang18.rds")
```

### UMAPs

```{r}
cbf_palette = c(
    "#E69F00",  # Orange
    "#56B4E9",  # Sky Blue
    "#009E73",  # Bluish Green
    "#F0E442",  # Yellow
    "#0072B2",  # Blue
    "#D55E00",  # Vermillion
    "#CC79A7",  # Reddish Purple
    "#999999",  # Gray
    "#117733",  # Dark Green
    "#882255",  # Wine
    "#44AA99",  # Teal
    "#DDCC77",  # Sand
    "#88CCEE",  # Light Blue
    "#AA4499",  # Magenta
    "#332288",  # Navy
    "#E17C05",  # Burnt Orange
    "#DC267F",  # Pink
    "#648FFF",  # Vivid Blue
    "#785EF0",  # Violet
    "#FE6100"   # Bright Orange
)
```

```{r}
# define selections to be used for UMAPs
selections_clu_con <- c(
  "random",
  "HVG",
  "tF",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPBDS_tF",
  "sPVE_tF",
  "sPVE_tPVE",
  "common_score",
  "tMAV",
  "tEN",
  "tMAD",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV",
  "sEN",
  "sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD",
  "all_genes"
)

selections_con <- c(
  "common_score",
  "sMAV",
  "sEN",
  "sAD",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)
```

```{r}
umap_tsne_plots <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  sce_iter <- sce
  
  # create UMAP plots
  plots <- c()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    sce_iter <- runUMAP(sce_iter, dimred="PCA")
    sce_iter <- runTSNE(sce_iter, dimred="PCA")
    
    # create UMAP plot colored by condition
    plot1 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "UMAP")),
      mapping = aes(x = UMAP1, y = UMAP2, colour = sce_iter$group_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[1:2]) +
      labs(title = "",
           x = "UMAP1",
           y = "UMAP2",
           colour = "Treatment") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 18,
            face = "bold"
          )
      )
    
    # create UMAP plot colored by cell type
    plot2 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "UMAP")),
      mapping = aes(x = UMAP1, y = UMAP2, colour = sce_iter$cluster_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[3:10]) +
      labs(title = selection,
           x = "",
           y = "UMAP2",
           colour = "Cell Type") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 20,
            face = "bold"
          )
      )
    
    # create TSNE plot colored by condition
    plot3 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "TSNE")),
      mapping = aes(x = TSNE1, y = TSNE2, colour = sce_iter$group_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[1:2]) +
      labs(title = "",
           x = "TSNE1",
           y = "TSNE2",
           colour = "Treatment") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 20,
            face = "bold"
          )
      )
    
    # create TSNE plot colored by cell type
    plot4 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "TSNE")),
      mapping = aes(x = TSNE1, y = TSNE2, colour = sce_iter$cluster_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[3:10]) +
      labs(title = selection,
           x = "",
           y = "TSNE2",
           colour = "Cell Type") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 20,
            face = "bold"
          )
      )
    
    plots <- c(
      plots,
      plot1,
      plot2,
      plot3,
      plot4
    )
    
  }
  
  return(plots)
  
})
```

```{r}
umap_tsne_plots$Kang18_clu_con
```

```{r}
for (i in 1:length(umap_tsne_plots$Kang18_clu_con)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_tsne_plots$Kang18_clu_con[i], path = "../../outs/kang/umap_clu_con/",
         width = 4, height = 4, dpi = 300)
}
```

```{r}
umap_tsne_plots$Kang18_con
```

```{r}
for (i in 1:length(umap_tsne_plots$Kang18_con)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_tsne_plots$Kang18_con[i], path = "../../outs/kang/umap_con/",
         width = 4, height = 4, dpi = 300)
}
```

### Perform Clustering Based on Selections (ARI Scores)

```{r}
# define selections to be used for clustering
selections_clu_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD"
)

selections_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)

all_sels <- unique(c(selections_clu_con, selections_con))
color_map <- setNames(qualitative_hcl(21, palette = "Dark 3"), all_sels)
```

```{r}
ari_scores <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  sce_iter <- sce
  
  # calculate ARI scores
  ari_scores <- c()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction and clustering
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    g <- buildSNNGraph(sce_iter, use.dimred="PCA", type="jaccard")
    k <- cluster_louvain(g, resolution=0.3)$membership
    n_clusters <- length(unique(k))
    message(selection, ": ", n_clusters, " clusters")
    new_col <- paste0("cluster_re_", selection)
    sce_iter[[new_col]] <- factor(k, seq_along(unique(k)))
    
    # ARI score calculation
    ari_scores <- c(
      ari_scores,
      ARI(sce_iter$cluster_id, sce_iter[[new_col]])
    )
    
  }
  
  results <- data.frame(Selection = selections, ARI = ari_scores)
  results$Selection <- factor(results$Selection, levels = results$Selection)
  
  return(results)
  
})
```

```{r}
p1 <- ari_scores$Kang18_clu_con %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Cell Types & Conditions")

p1
```

```{r}
p2 <- ari_scores$Kang18_con %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Conditions")

p2
```

```{r}
# combine individual plots
ari_plots <- list(p1, p2)

ari_plots <- lapply(ari_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(ari_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "ARI Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_ari.png", combined_plot, path = "../../outs/kang/",
       width = 12, height = 6, dpi = 300)
combined_plot
```


















