---
title: "Scores-Selections Comparison"
author: "Michel Tarnow"
date: today
format: 
    html:
      toc: true
      self-contained: true
      highlight-style: github
      code-line-numbers: true
editor_options: 
    chunk_output_type: inline
---

## Import Packages

```{r}
suppressPackageStartupMessages({
  library(reshape2)
  library(MLmetrics)
  library(caret)
  library(ComplexUpset)
  library(patchwork)
  library(scran)
  library(scater)
  library(igraph)
  library(SingleCellExperiment)
  library(zellkonverter)
  library(MLVSBM)
  library(tidyverse)
  library(viridis)
  library(GGally)
  library(rlang)
  library(colorspace)
  library(lisi)
})
```



## Scores

Patches scores produced by running the model on simulation data in Python are loaded from file and recombined into single group and condition scores. For example, scores for the different conditions (one score per condition) are combined into a single score by their max absolute value or by their Euclidean norm. FSS scores, which were obtained by running the snakemake workflow, are loaded from file. The different scores are finally compared.

### Helper Functions

```{r}
# helper function for loading and processing Patches scores
load_process_patches_scores <- function(file) {
  
  # read in df from file and fix column names
  df <- read_csv(
    file,
    col_names = TRUE,
    name_repair = "minimal",
    show_col_types = FALSE
  )
  colnames(df)[1] = "gene_id"
  
  # remove _Patches from column names
  df <- df %>%
    rename_with(~ sub("_Patches$", "", .x), ends_with("_Patches"))
  
  # add combinations of scores
  
  ## max absolute value for condition
  df <- df %>% 
    mutate(
      sMAV = pmax(
        abs(Condition1_score),
        abs(Condition2_score)
      )
    )
  
  ## Euclidean norm for condition
  df <- df %>% 
    mutate(
      sEN = sqrt(
        Condition1_score^2 +
          Condition2_score^2
      )
    )
  
  ## absolute difference for condition
  df <- df %>% 
    mutate(
      sAD = abs(Condition1_score - Condition2_score))
    
  if ("Group1_score" %in% names(df)) {
    # cluster and condition scores available
    
    ## max absolute value for group
    df <- df %>% 
      mutate(
        tMAV = pmax(
          abs(Group1_score),
          abs(Group2_score),
          abs(Group3_score)
        )
      )
    
    ## Euclidean norm for group
    df <- df %>% 
      mutate(
        tEN = sqrt(
          Group1_score^2 +
            Group2_score^2 +
            Group3_score^2
        )
      )
    
    ## max absolute difference for group
    # df <- df %>% 
      # mutate(
        # tXAD = pmax(
          # abs(Group1_score - Group2_score),
          # abs(Group1_score - Group3_score),
          # abs(Group2_score - Group3_score)
        # )
      # )
    
    ## mean absolute difference for group
    df <- df %>% 
      mutate(
        tMAD = rowMeans(
          cbind(
            abs(Group1_score - Group2_score),
            abs(Group1_score - Group3_score),
            abs(Group2_score - Group3_score)
          )
        )
      )
    
    ## sum of absolute differences for group
    # df <- df %>% 
      # mutate(
        # tSAD = rowSums(
          # cbind(
            # abs(Group1_score - Group2_score),
            # abs(Group1_score - Group3_score),
            # abs(Group2_score - Group3_score)
          # )
        # )
      #)
    
  }
  
  return(df)
    
}

# helper function for loading and processing FSS scores
load_process_fss_scores <- function(files, hvgs) {
  
  # select gene and score columns and combine all dfs in one list
  dfs <- lapply(files, function(file) {
    
    # read in df from file
    df <- readRDS(file)
    
    # rename score value column to score name
    if (nrow(df) > length(unique(df$gene_id))) {
      # df contains both tPVE and sPVE scores
      
      df <- df %>% 
        pivot_wider(names_from = sco, values_from = sco_val)
      cols_to_keep <- c("gene_id", "tPVE", "sPVE")
      
    } else {
      # df contains a single score (all but tPVE + sPVE)
      
      new_colname <- df$sco[1]
      df <- df %>% 
        rename(!!new_colname := sco_val)
      cols_to_keep <- c("gene_id", new_colname)
      
    }
    
    # select gene and score columns
    df <- df %>% 
      select(all_of(cols_to_keep))
    
    # return df
    return(df)
    
  })
  
  # combine all dfs in one df
  df_combined <- as_tibble(reduce(
    dfs, 
    ~ inner_join(.x, .y, by = "gene_id")
  ))
  
  # create df HVGs only
  df_combined_hvg <- df_combined %>% 
    filter(gene_id %in% hvgs)
  
  # return combined df
  return(
    list(
      fss_full = df_combined,
      fss_hvg = df_combined_hvg
    )
  )
  
}
```

### Load and Process Patches Scores

```{r}
files <- list.files(
  path = "../../data/sim/02-scores_patches/",
  pattern = "\\.csv$", full.names = TRUE,
)

patches_scores <- lapply(files, load_process_patches_scores)
names(patches_scores) <- sub(".*-(.*)_loadings\\.csv", "\\1", basename(files))
```

```{r}
lapply(patches_scores, head)
```

### Load and Process FSS Scores

```{r}
files <- list.files(
  "../../data/sim/02-scores_fss/",
  pattern = "\\.rds$", 
  full.names = TRUE
)
hvgs <- patches_scores$patches_clu_con_hvg$gene_id

fss_scores <- load_process_fss_scores(files, hvgs)
```

```{r}
lapply(fss_scores, head)
```

### Combine and Transform Scores

```{r}
# combine Patches and FSS scores in one df for each Patches workflow
# and add z-transformed scores
# and add ranks
all_scores <- lapply(patches_scores, function(df) {
  
  if (dim(df)[1] == dim(fss_scores$fss_full)[1]) {
    # df contains full data
    
    new_df <- cbind.data.frame(
      df,
      fss_scores$fss_full %>% 
        select(!gene_id)
    )
    
  } else {
    # df contains HVG data
    
    new_df <- cbind.data.frame(
      df,
      fss_scores$fss_hvg %>% 
        select(!gene_id)
    )
    
  }
  
  # add z-transformed scores and ranks
  if ("common_score_z" %in% colnames(df)) {
    # z-transformations already exist, skip df
    return(new_df)
  }
  
  new_df <- new_df %>% 
    mutate(
      across(
        .cols = -gene_id,
        .fns = list(
          z = ~ as.numeric(scale(.x)),
          rank = ~ rank(-.x, ties.method = "first")
        ),
        .names = "{.col}_{.fn}"
      )
    )
  
  return(new_df)
  
})

lapply(all_scores, head)
```

### Compare Scores

#### Correlation Between Scores

```{r}
# create correlation plots
cor_plots_rank <- lapply(all_scores, function(df) {
  
  # select ranked scores
  df_rank <- df %>% 
    select(ends_with("_rank"))
  
  # calculate pairwise Spearman rank correlations
  cor_mat <- cor(
    df_rank,
    use = "pairwise.complete.obs",
    method = "spearman"
  )
  
  # plot correlations as a heat map
  cor_long <- melt(cor_mat)
  ggplot(data = cor_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(-1, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Correlation")
  
})

cor_plots_rank
```

```{r}
# combine individual plots
subtitles <- c(
  "Full Data - Groups & Conditions",
  "HVG Data - Groups & Conditions",
  "Full Data - Conditions",
  "HVG Data - Conditions"
)
cor_plots_rank <- Map(function(p, sub) p + ggtitle(sub), cor_plots_rank, subtitles)

cor_plots_rank <- lapply(cor_plots_rank, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(cor_plots_rank, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Ranked Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_rank.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 10, dpi = 300)
combined_plot
```

```{r}
# create correlation plots
cor_plots_z <- lapply(all_scores, function(df) {
  
  # select z-transformed scores
  df_z <- df %>% 
    select(ends_with("_z"))
  
  # calculate pairwise Spearman rank correlations
  cor_mat <- cor(
    df_z,
    use = "pairwise.complete.obs",
    method = "spearman"
  )
  
  # plot correlations as a heat map
  cor_long <- melt(cor_mat)
  ggplot(data = cor_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(-1, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Correlation")
  
})

cor_plots_z
```

```{r}
# combine individual plots
subtitles <- c(
  "Full Data - Groups & Conditions",
  "HVG Data - Groups & Conditions",
  "Full Data - Conditions",
  "HVG Data - Conditions"
)
cor_plots_z <- Map(function(p, sub) p + ggtitle(sub), cor_plots_z, subtitles)

cor_plots_z <- lapply(cor_plots_z, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(cor_plots_z, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Z-Transformed Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_z.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 10, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)
plots <- cor_plots_z[c(1, 3)]
plots <- Map(function(p, sub) p + ggtitle(sub), plots, subtitles)

plots <- lapply(plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Z-Transformed Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_z_full.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 6, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)
plots <- cor_plots_z[c(2, 4)]
plots <- Map(function(p, sub) p + ggtitle(sub), plots, subtitles)

plots <- lapply(plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Spearman Correlation of Z-Transformed Scores",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_cor_z_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 6, dpi = 300)
combined_plot
```

#### Overlap of top-n genes (Jaccard Index)

```{r}
# function to compute the top-n genes for a given rank
get_top_n_genes <- function(df, rank, n) {
  top_genes <- (df %>% 
    arrange(.data[[rank]]) %>% 
    slice_head(n = n))$gene_id
  
  return(top_genes)
}

# function to compute Jaccard index for top n genes
compute_jaccard <- function(df, rank1, rank2, n) {
  top1 <- get_top_n_genes(df, rank1, n)
  top2 <- get_top_n_genes(df, rank2, n)
  length(intersect(top1, top2)) / length(union(top1, top2))
}

# function to compute Jaccard matrix for all provided ranks
compute_jaccard_matrix <- function(df, n) {
  
  rank_cols <- grep("_rank$", colnames(df), value = TRUE)
  m <- length(rank_cols)
  
  jmat <- matrix(NA, nrow = m, ncol = m, dimnames = list(rank_cols, rank_cols))
  
  for (i in seq_along(rank_cols)) {
    for (j in seq_along(rank_cols)) {
      jmat[i, j] <- compute_jaccard(df, rank_cols[i], rank_cols[j], n)
    }
  }
  
  return(jmat)
  
}
```

```{r}
# plot Jaccard indices as heat maps
j_plots <- lapply(all_scores, function(df) {
  
  j_mat <- compute_jaccard_matrix(df, 100)
  
  j_long <- melt(j_mat)
  ggplot(data = j_long, mapping = aes(Var1, Var2, fill = value)) +
    geom_tile() +
    scale_fill_viridis_c(
      option = "C",
      direction = 1,
      limits = c(0, 1)
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank(),
      plot.title = element_text(size = 11)
    ) +
    labs(x = "", y = "", fill = "Jaccard\nIndex")
  
})

j_plots
```

```{r}
# combine individual plots
subtitles <- c(
  "Full Data - Groups & Conditions",
  "HVG Data - Groups & Conditions",
  "Full Data - Conditions",
  "HVG Data - Conditions"
)
j_plots <- Map(function(p, sub) p + ggtitle(sub), j_plots, subtitles)

j_plots <- lapply(j_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(j_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Jaccard Index of Top-100 Genes",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_jaccard.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 10, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)

plots <- j_plots[c(1, 3)]
plots <- Map(function(p, sub) p + ggtitle(sub), plots, subtitles)

plots <- lapply(plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Jaccard Index of Top-100 Genes",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_jaccard_full.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 6, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)

plots <- j_plots[c(2, 4)]
plots <- Map(function(p, sub) p + ggtitle(sub), plots, subtitles)

plots <- lapply(plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Jaccard Index of Top-100 Genes",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sco_jaccard_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 6, dpi = 300)
combined_plot
```



## Selections

### Load and Process Gene Ground Truths (FSS Selections)

Gene ground truths, i.e., whether a gene is considered a type or a state gene, can be obtained from the simulated data via the logFC of splatter's DE factor (e.g., `ConditionDE.Condition1` or `GroupDE.Group1` in gene metadata). A gene is considered DEnotDS if there is a group logFC unequal to 1 and no condition logFC unequal to 1. Similarly, a gene is considered DS if there is a condition logFC unequal to 1, regardless of group logFCs. DEnotDS genes are selected as ground truth type genes, while DS genes are selected for ground truth state genes, i.e., a type gene must have a group effect and no state effect, while a state gene must have a condition effect while a group effect is not necessary (but can exist).

Here, all selections from the FSS paper's workflow are loaded for the t100,s80,b0 experiment and combined. This includes the selection strategies based on simulation parameters (like DEnotDS and DS) as well as selection strategies based on feature scores. 

```{r}
load_process_fss_selections <- function(files, hvgs) {
  
  # select gene id and selection columns and combine all dfs in one list
  fss_selection_dfs <- lapply(files, function(file) {
    
    # read in df from file
    df <- readRDS(file)
    
    # df contains a single selection
    new_colname <- df$sel[1]
    df <- df %>% 
      rename(!!new_colname := sel_val)
    cols_to_keep <- c("gene_id", new_colname)
    
    # select gene and score columns
    df <- df %>% 
      select(all_of(cols_to_keep))
    
    return(df)
    
  })
  
  # combine all dfs in one df
  fss_selections_full <- as_tibble(reduce(
    fss_selection_dfs, 
    ~ inner_join(.x, .y, by = "gene_id")
  ))
  
  # sort columns to put ground truths first
  fss_selections_full <- fss_selections_full %>% 
    select(gene_id, DEnotDS, DS, DSnotDE, DEgtDS, DSgtDE, DE, everything())
  
  # create df with only HVGs
  fss_selections_hvg <- fss_selections_full %>% 
    filter(gene_id %in% hvgs)
  
  # combine into list
  fss_selections <- list(
    fss_full = fss_selections_full,
    fss_hvg = fss_selections_hvg
  )
  
  return(fss_selections)
  
}
```

```{r}
# get all files containing selections
rds_files <- list.files(
  "../../data/sim/03-selections_fss/",
  pattern = "\\.rds$", 
  full.names = TRUE
)

# get HVGs
hvgs <- patches_scores$patches_clu_con_hvg$gene_id

# load and process FSS selections
fss_selections <- load_process_fss_selections(rds_files, hvgs)

lapply(fss_selections, head)
```

```{r}
# get number of genes to select from ground truth (match the number of ground
# truth type (DEnotDS) and state (DS) genes for full and HVG data)
nt_full <- sum(fss_selections$fss_full$DEnotDS)
nt_hvg <- sum(fss_selections$fss_hvg$DEnotDS)
ns_full <- sum(fss_selections$fss_full$DS)
ns_hvg <- sum(fss_selections$fss_hvg$DS)

# number of genes for DSnotDE
nst_full <- sum(fss_selections$fss_full$DSnotDE)
nst_hvg <- sum(fss_selections$fss_hvg$DSnotDE)
```

```{r}
# add new FSS selections
fss_selections <- lapply(fss_selections, function(df) {
  
  # determine if df contains full or HVG data and adjust n accordingly
  full <- ifelse(dim(df)[1] > 1500, TRUE, FALSE)
  nt <- ifelse(full, nt_full, nt_hvg)
  ns <- ifelse(full, ns_full, ns_hvg)
  nst <- ifelse(full, nst_full, nst_hvg)
  
  # determine if df contains full or HVG data and adjust score df accordingly
  score_df <- if (full) {
    fss_scores$fss_full
  } else {
    fss_scores$fss_hvg
  }
  
  # tF and sPBDS
  score_names <- c("tF", "sPBDS")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    df$sPBDS_tF <- scores >= cutoff
    
  }
  
  # tF and sPVE
  score_names <- c("tF", "sPVE")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    df$sPVE_tF <- scores >= cutoff
    
  }
  
  # tPVE and sPVE
  score_names <- c("tPVE", "sPVE")
  if (all(score_names %in% colnames(score_df))) {
    
    t <- rank(score_df[[score_names[1]]])
    s <- rank(score_df[[score_names[2]]])
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    df$sPVE_tPVE <- scores >= cutoff
    
  }
  
  return(df)
  
})
```

### Compute Patches Selections

```{r}
patches_selections <- lapply(patches_scores, function(score_df) {
  
  # determine if df contains full or HVG data and adjust n accordingly
  full <- ifelse(dim(score_df)[1] > 1500, TRUE, FALSE)
  nt <- ifelse(full, nt_full, nt_hvg)
  ns <- ifelse(full, ns_full, ns_hvg)
  nst <- ifelse(full, nst_full, nst_hvg)
  
  # create binary df that contains binary selections based on scores
  binary_df <- score_df
  for (col in names(score_df)) {
    
    # determine whether score is a type or state score and adjust n accordingly 
    if (grepl("Group|group|common|tMAV|tEN|tMAD", col)) {
      # score is a type score
      n <- nt
    } else if (grepl("condition|Condition|sMAV|sEN|sAD", col)) {
      # score is a state score
      n <- ns
    } else {
      # skip columns not matching rules
      next
    }
    
    # convert score into binary predictions
    scores <- score_df[[col]]
    cutoff <- sort(scores, decreasing = TRUE)[n]
    binary_df[[col]] <- scores >= cutoff
    
  }
  
  # add new selections based on combinations of scores
  
  # difference in rank of max absolute group and condition scores
  max_abs <- c("tMAV", "sMAV")
  if (all(max_abs %in% colnames(score_df))) {
    
    t <- rank(score_df[[max_abs[1]]])
    s <- rank(score_df[[max_abs[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$tMAV_sMAV <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sMAV_tMAV <- scores >= cutoff
    
  }
  
  # difference in rank of Euclidean norm group and condition scores
  eucl_norm <- c("tEN", "sEN")
  if (all(eucl_norm %in% colnames(score_df))) {
    
    t <- rank(score_df[[eucl_norm[1]]])
    s <- rank(score_df[[eucl_norm[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$tEN_sEN <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sEN_tEN <- scores >= cutoff
    
  }
  
  # difference in rank of absolute difference group and condition scores
  abs_dif <- c("tMAD", "sAD")
  if (all(abs_dif %in% colnames(score_df))) {
    
    t <- rank(score_df[[abs_dif[1]]])
    s <- rank(score_df[[abs_dif[2]]])
    
    # type not state
    scores <- t - s
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$tMAD_sAD <- scores >= cutoff
    
    # state not type
    scores <- s - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sAD_tMAD <- scores >= cutoff
    
  }
  
  # difference in rank of common score and condition scores
  com_con <- c("common_score", "sMAV", "sEN", "sAD")
  if (all(com_con %in% colnames(score_df)) && !("Group1_score" %in% colnames(score_df))) {
    
    t <- rank(score_df[[com_con[1]]])
    s1 <- rank(score_df[[com_con[2]]])
    s2 <- rank(score_df[[com_con[3]]])
    s3 <- rank(score_df[[com_con[4]]])
    
    # common - sMAV
    scores <- t - s1
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$common_sMAV <- scores >= cutoff
    
    # common - sEN
    scores <- t - s2
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$common_sEN <- scores >= cutoff
    
    # common - sAD
    scores <- t - s3
    cutoff <- sort(scores, decreasing = TRUE)[nt]
    binary_df$common_sAD <- scores >= cutoff
    
    # sMAV - common
    scores <- s1 - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sMAV_common <- scores >= cutoff
    
    # sEN - common
    scores <- s2 - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sEN_common <- scores >= cutoff
    
    # sAD - common
    scores <- s3 - t
    cutoff <- sort(scores, decreasing = TRUE)[ns]
    binary_df$sAD_common <- scores >= cutoff
    
  }
  
  # remove selections not needed for downstream analyses
  if ("Group1_score" %in% colnames(binary_df)) {
    binary_df <- binary_df %>% 
      select(
        -Group1_score,
        -Group2_score,
        -Group3_score
      )
  }
  binary_df <- binary_df %>% 
    select(
      -Condition1_score,
      -Condition2_score
    )
  
  return(binary_df)
  
})

names(patches_selections) <- names(patches_scores)
lapply(patches_selections, head)
```

### Combine Selections

```{r}
# combine Patches and FSS selections in one df for each Patches workflow
all_selections <- lapply(patches_selections, function(df) {
  
  if (dim(df)[1] == dim(fss_selections$fss_full)[1]) {
    # df contains full data
    
    new_df <- cbind.data.frame(
      df,
      fss_selections$fss_full %>% 
        select(!gene_id)
    )
    
  } else {
    # df contains HVG data
    
    new_df <- cbind.data.frame(
      df,
      fss_selections$fss_hvg %>% 
        select(!gene_id)
    )
    
  }
  
  # add all selection as baseline for downstream analyses
  new_df$all_genes <- TRUE
  
  return(new_df)
  
})

lapply(all_selections, head)
```

### Compare Selections

#### Overlaps

```{r}
plot_overlaps <- function(df, title) {
  
  # remove columns and scores not needed
  df <- df %>% 
    select(
      -DEgtDS,
      -DSgtDE,
      -DE
    )
  
  # define ground truths as reference columns
  ref_cols <- c("DEnotDS", "DS", "DSnotDE")
  
  # define scores to compare to ground truth
  score_cols <- setdiff(colnames(df), ref_cols)

  # compute overlaps
  overlap_df <- expand.grid(Score = score_cols, GroundTruth = ref_cols, stringsAsFactors = FALSE) %>% 
    rowwise() %>% 
    mutate(Overlap = sum(df[[Score]] & df[[GroundTruth]])) %>% 
    ungroup()
  
  # order scores for plot
  overlap_df$Score <- factor(overlap_df$Score, levels = score_cols)
  
  # compute number of TRUEs per GroundTruth
  counts <- colSums(df[ref_cols])
  overlap_df$GroundTruth <- factor(overlap_df$GroundTruth, 
                                   levels = ref_cols,
                                   labels = paste0(ref_cols, " (", counts, ")"))
  
  # plot overlap as heat map
  ggplot(overlap_df, aes(x = Score, y = GroundTruth, fill = Overlap)) +
    geom_tile(color = "white") +
    geom_text(aes(label = Overlap), color = "black") +
    scale_fill_viridis_c(option = "C", direction = -1) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = title, x = "", y = "")
  
}
```

```{r}
p1 <- plot_overlaps(
  bind_cols(
    fss_selections$fss_full %>% select(-gene_id),
    patches_selections$patches_clu_con_full %>% select(-gene_id)
  ),
  "Full Data - Groups & Conditions"
)
p1
```

```{r}
p2 <- plot_overlaps(
  bind_cols(
    fss_selections$fss_hvg %>% select(-gene_id),
    patches_selections$patches_clu_con_hvg %>% select(-gene_id)
  ),
  "HVG Data - Groups & Conditions"
)
p2
```

```{r}
p3 <- plot_overlaps(
  bind_cols(
    fss_selections$fss_full %>% select(-gene_id),
    patches_selections$patches_con_full %>% select(-gene_id)
  ),
  "Full Data - Conditions"
)
p3
```

```{r}
p4 <- plot_overlaps(
  bind_cols(
    fss_selections$fss_hvg %>% select(-gene_id),
    patches_selections$patches_con_hvg %>% select(-gene_id)
  ),
  "HVG Data - Conditions"
)
p4
```

```{r}
# combine individual plots
overlap_plots <- list(p1, p2, p3, p4)
subtitles <- c(
  "Full Data - Groups & Conditions",
  "HVG Data - Groups & Conditions",
  "Full Data - Conditions",
  "HVG Data - Conditions"
)
overlap_plots <- Map(function(p, sub) p + ggtitle(sub), overlap_plots, subtitles)

overlap_plots <- lapply(overlap_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(overlap_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "Overlap of Selected Genes",
    theme = theme(
      legend.position = "none",
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_overlap.png", combined_plot, path = "../../outs/sim/",
       width = 14, height = 9, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
overlap_plots <- list(p1, p3)
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)
overlap_plots <- Map(function(p, sub) p + ggtitle(sub), overlap_plots, subtitles)

overlap_plots <- lapply(overlap_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(overlap_plots, ncol = 1, guides = "collect") +
  plot_annotation(
    title = "Overlap of Selected Genes",
    theme = theme(
      legend.position = "none",
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_overlap_full.png", combined_plot, path = "../../outs/sim/",
       width = 9, height = 8, dpi = 300)
combined_plot
```

```{r}
# combine individual plots
overlap_plots <- list(p2, p4)
subtitles <- c(
  "Groups & Conditions",
  "Conditions"
)
overlap_plots <- Map(function(p, sub) p + ggtitle(sub), overlap_plots, subtitles)

overlap_plots <- lapply(overlap_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(overlap_plots, ncol = 1, guides = "collect") +
  plot_annotation(
    title = "Overlap of Selected Genes",
    theme = theme(
      legend.position = "none",
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_overlap_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 9, height = 8, dpi = 300)
combined_plot
```

#### F1 Scores

```{r}
# define selections to be used for F1 score calculations
selections_clu_con <- c(
  "random",
  "HVG",
  "tF",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_score",
  "sMAV",
  "sEN",
  "sAD",
  "tMAV",
  "tEN",
  "tMAD",
  "tMAV_sMAV",
  "sMAV_tMAV",
  "tEN_sEN",
  "sEN_tEN",
  "tMAD_sAD",
  "sAD_tMAD"
)

selections_con <- c(
  "random",
  "HVG",
  "tF",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_score",
  "sMAV",
  "sEN",
  "sAD",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)
```

```{r}
f1_scores <- lapply(all_selections, function(df) {
  
  # get ground truth
  y_true_type <- df$DEnotDS
  y_true_state <- df$DS
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  # calculate F1 scores
  f1_scores <- c()
  for (selection in selections) {

    if (str_starts(selection, "s")) {
      # state scores: compare to y_true_state
      f1_scores <- c(
        f1_scores,
        confusionMatrix(
          factor(y_true_state, levels = c(FALSE, TRUE)),
          factor(df[[selection]], levels = c(FALSE, TRUE)),
          positive = "TRUE"
        )$byClass["F1"]
      )
      
    } else {
      # type scores: compare to y_true_type
      f1_scores <- c(
        f1_scores,
        confusionMatrix(
          factor(y_true_type, levels = c(FALSE, TRUE)),
          factor(df[[selection]], levels = c(FALSE, TRUE)),
          positive = "TRUE"
        )$byClass["F1"]
      )
      
    }
    
  }
  
  results <- data.frame(Selection = selections, F1 = f1_scores)
  results$Selection <- factor(results$Selection, levels = results$Selection)
  
  return(results)
  
})
```

```{r}
p1 <- bind_cols(
  f1_scores$patches_clu_con_full,
  f1_scores$patches_clu_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = c(Full, HVG), names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = Data)) +
    geom_col(position = position_dodge(width = 0.8)) +
    coord_flip() +
    scale_fill_manual(values = c("steelblue", "darkgrey")) +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Groups & Conditions")

p1
```

```{r}
p2 <- bind_cols(
  f1_scores$patches_con_full,
  f1_scores$patches_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = c(Full, HVG), names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = Data)) +
    geom_col(position = position_dodge(width = 0.8)) +
    coord_flip() +
    scale_fill_manual(values = c("steelblue", "darkgrey")) +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Conditions")

p2
```

```{r}
# combine individual plots
f1_plots <- list(p1, p2)

f1_plots <- lapply(f1_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(f1_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "F1 Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_f1.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 5, dpi = 300)
combined_plot
```

```{r}
p3 <- bind_cols(
  f1_scores$patches_clu_con_full,
  f1_scores$patches_clu_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = Full, names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = I("steelblue"))) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Groups & Conditions")

p3
```

```{r}
p4 <- bind_cols(
  f1_scores$patches_con_full,
  f1_scores$patches_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = Full, names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = I("steelblue"))) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Conditions")

p4
```

```{r}
# combine individual plots
f1_plots <- list(p3, p4)

f1_plots <- lapply(f1_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(f1_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "F1 Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_f1_full.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 5, dpi = 300)
combined_plot
```

```{r}
p5 <- bind_cols(
  f1_scores$patches_clu_con_full,
  f1_scores$patches_clu_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = HVG, names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = I("steelblue"))) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Groups & Conditions")

p5
```

```{r}
p6 <- bind_cols(
  f1_scores$patches_con_full,
  f1_scores$patches_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = F1...2, HVG = F1...3) %>% 
  pivot_longer(cols = HVG, names_to = "Data", values_to = "F1") %>% 
  ggplot(aes(x = reorder(Selection, F1), y = F1, fill = I("steelblue"))) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    labs(x = "", y = "F1", title = "Conditions")

p6
```

```{r}
# combine individual plots
f1_plots <- list(p5, p6)

f1_plots <- lapply(f1_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(f1_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "F1 Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_f1_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 6, dpi = 300)
combined_plot
```

### Export Selections

```{r}
# read in sce file
sce_full <- readRDS("../../data/sim/01-pro/t100,s80,b0.rds")

# create HVG sce object
hvgs <- fss_selections$fss_hvg$gene_id
sce_hvg <- sce_full[rownames(sce_full) %in% hvgs, ]

# function for adding selection to sce objects
add_selection_data <- function(sce, selections_list, name) {
  
  selection <- selections_list[[name]]
  if (!all(rownames(sce) == selection$gene_id)) {
    stop("Gene IDs do not match between SCE and selection table")
  }
  
  rowData(sce) <- cbind(rowData(sce), selection[ , -1])
  return(sce)
  
}

# add selections to sce objects
sce_clu_con_full <- add_selection_data(sce_full, all_selections, "patches_clu_con_full")
sce_con_full     <- add_selection_data(sce_full, all_selections, "patches_con_full")
sce_clu_con_hvg  <- add_selection_data(sce_hvg,  all_selections, "patches_clu_con_hvg")
sce_con_hvg      <- add_selection_data(sce_hvg,  all_selections, "patches_con_hvg")

# export sce to anndata
writeH5AD(sce_clu_con_full, "../../data/sim/03-selections/t100,s80,b0-selections_clu_con_full.h5ad")
writeH5AD(sce_con_full, "../../data/sim/03-selections/t100,s80,b0-selections_con_full.h5ad")
writeH5AD(sce_clu_con_hvg, "../../data/sim/03-selections/t100,s80,b0-selections_clu_con_hvg.h5ad")
writeH5AD(sce_con_hvg, "../../data/sim/03-selections/t100,s80,b0-selections_con_hvg.h5ad")
```



## Downstream Analyses

### Load SingleCellExperiment

```{r}
# full data
sce_full <- readRDS("../../data/sim/01-pro/t100,s80,b0.rds")

# HVG data
hvgs <- fss_selections$fss_hvg$gene_id
sce_hvg <- sce_full[rownames(sce_full) %in% hvgs, ]
```

### UMAPs

```{r}
cbf_palette = c(
    "#E69F00",  # Orange
    "#56B4E9",  # Sky Blue
    "#009E73",  # Bluish Green
    "#F0E442",  # Yellow
    "#0072B2",  # Blue
    "#D55E00",  # Vermillion
    "#CC79A7",  # Reddish Purple
    "#999999",  # Gray
    "#117733",  # Dark Green
    "#882255",  # Wine
    "#44AA99",  # Teal
    "#DDCC77",  # Sand
    "#88CCEE",  # Light Blue
    "#AA4499",  # Magenta
    "#332288",  # Navy
    "#E17C05",  # Burnt Orange
    "#DC267F",  # Pink
    "#648FFF",  # Vivid Blue
    "#785EF0",  # Violet
    "#FE6100"   # Bright Orange
)
```

```{r}
# define selections to be used for UMAPs
selections_clu_con <- c(
  "random",
  "HVG",
  "tF",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_score",
  "tMAV",
  "tEN",
  "tMAD",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV",
  "sEN",
  "sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD",
  "all_genes"
)

selections_con <- c(
  "common_score",
  "sMAV",
  "sEN",
  "sAD",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)
```

```{r}
umap_plots <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  # determine if df contains full or HVG data
  if (dim(df)[1] == dim(sce_full)[1]) {
    # df contains full data
    sce_iter <- sce_full
  } else {
    # df contains HVG data
    sce_iter <- sce_hvg
  }
  
  # create UMAP plots
  plots <- c()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    sce_iter <- runUMAP(sce_iter, dimred="PCA")

    # create UMAP plot colored by condition
    plot1 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "UMAP")),
      mapping = aes(x = UMAP1, y = UMAP2, colour = sce_iter$group_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[1:2]) +
      labs(title = "",
           x = "UMAP1",
           y = "UMAP2",
           colour = "Condition") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 20,
            face = "bold"
          )
      )
    
    # create UMAP plot colored by group
    plot2 <- ggplot(
      data = as.data.frame(reducedDim(sce_iter, "UMAP")),
      mapping = aes(x = UMAP1, y = UMAP2, colour = sce_iter$cluster_id)
    ) +
      geom_point(size = 1, alpha = 0.7) +
      scale_colour_manual(values = cbf_palette[c(3, 5, 7)]) +
      labs(title = selection,
           x = "",
           y = "UMAP2",
           colour = "Group") +
      theme_classic() +
      theme(
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        legend.position = "none",
        text = element_text(size = 16),
        plot.title = element_text(
            hjust = 0.5,
            size = 20,
            face = "bold"
          )
      )
    
    plots <- c(
      plots,
      plot1,
      plot2
    )
    
  }
  
  return(plots)
  
})
```

```{r}
umap_plots$patches_clu_con_full
```

```{r}
for (i in 1:length(umap_plots$patches_clu_con_full)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_plots$patches_clu_con_full[i], path = "../../outs/sim/umap_clu_con_full/",
         width = 4, height = 4, dpi = 300)
}
```

```{r}
umap_plots$patches_clu_con_hvg
```

```{r}
for (i in 1:length(umap_plots$patches_clu_con_hvg)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_plots$patches_clu_con_hvg[i], path = "../../outs/sim/umap_clu_con_hvg/",
         width = 4, height = 4, dpi = 300)
}
```

```{r}
umap_plots$patches_con_full
```

```{r}
for (i in 1:length(umap_plots$patches_con_full)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_plots$patches_con_full[i], path = "../../outs/sim/umap_con_full/",
         width = 4, height = 4, dpi = 300)
}
```

```{r}
umap_plots$patches_con_hvg
```

```{r}
for (i in 1:length(umap_plots$patches_con_hvg)) {
  file_name <- paste0("umap_", i, ".png")
  
  ggsave(file_name, umap_plots$patches_con_hvg[i], path = "../../outs/sim/umap_con_hvg/",
         width = 4, height = 4, dpi = 300)
}
```

### ARI

```{r}
# define selections to be used for clustering
selections_clu_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD"
)

selections_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)

all_sels <- unique(c(selections_clu_con, selections_con))
color_map <- setNames(qualitative_hcl(21, palette = "Dark 3"), all_sels)
```

```{r}
ari_scores <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  # determine if df contains full or HVG data
  if (dim(df)[1] == dim(sce_full)[1]) {
    # df contains full data
    sce_iter <- sce_full
  } else {
    # df contains HVG data
    sce_iter <- sce_hvg
  }
  
  # calculate ARI scores
  ari_scores <- c()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction and clustering
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    g <- buildSNNGraph(sce_iter, use.dimred="PCA", type="jaccard")
    k <- cluster_louvain(g, resolution=1)$membership
    n_clusters <- length(unique(k))
    message(selection, ": ", n_clusters, " clusters")
    new_col <- paste0("cluster_re_", selection)
    sce_iter[[new_col]] <- factor(k, seq_along(unique(k)))
    
    # ARI score calculation
    ari_scores <- c(
      ari_scores,
      ARI(sce_iter$cluster_id, sce_iter[[new_col]])
    )
    
  }
  
  results <- data.frame(Selection = selections, ARI = ari_scores)
  results$Selection <- factor(results$Selection, levels = results$Selection)
  
  return(results)
  
})
```

```{r}
p1 <- bind_cols(
  ari_scores$patches_clu_con_full,
  ari_scores$patches_clu_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = ARI...2, HVG = ARI...3) %>% 
  pivot_longer(cols = c(Full, HVG), names_to = "Data", values_to = "ARI") %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Data)) +
    geom_col(position = position_dodge(width = 0.8)) +
    coord_flip() +
    scale_fill_manual(values = c("steelblue", "darkgrey")) +
    theme_minimal() +
    labs(x = "", y = "ARI", title = "Groups & Conditions")

p1
```

```{r}
p2 <- bind_cols(
  ari_scores$patches_con_full,
  ari_scores$patches_con_hvg %>% select(-Selection)
) %>% 
  rename(Full = ARI...2, HVG = ARI...3) %>% 
  pivot_longer(cols = c(Full, HVG), names_to = "Data", values_to = "ARI") %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Data)) +
    geom_col(position = position_dodge(width = 0.8)) +
    coord_flip() +
    scale_fill_manual(values = c("steelblue", "darkgrey")) +
    theme_minimal() +
    labs(x = "", y = "ARI", title = "Conditions")

p2
```

```{r}
# combine individual plots
ari_plots <- list(p1, p2)

ari_plots <- lapply(ari_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(ari_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "ARI Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_ari.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 5, dpi = 300)
combined_plot
```

```{r}
p3 <- ari_scores$patches_clu_con_full %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Groups & Conditions")

p3
```

```{r}
p4 <- ari_scores$patches_con_full %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Conditions")

p4
```

```{r}
# combine individual plots
ari_plots <- list(p3, p4)

ari_plots <- lapply(ari_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(ari_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "ARI Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_ari_full.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 5, dpi = 300)
combined_plot
```

```{r}
p5 <- ari_scores$patches_clu_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Groups & Conditions")

p5
```

```{r}
p6 <- ari_scores$patches_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, ARI), y = ARI, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "ARI", title = "Conditions")

p6
```

```{r}
# combine individual plots
ari_plots <- list(p5, p6)

ari_plots <- lapply(ari_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(ari_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "ARI Score per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_ari_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 6, dpi = 300)
combined_plot
```

### LISI

```{r}
# define selections to be used for clustering
selections_clu_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD"
)

selections_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)

all_sels <- unique(c(selections_clu_con, selections_con))
color_map <- setNames(qualitative_hcl(21, palette = "Dark 3"), all_sels)
```

```{r}
lisi_scores <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  # determine if df contains full or HVG data
  if (dim(df)[1] == dim(sce_full)[1]) {
    # df contains full data
    sce_iter <- sce_full
  } else {
    # df contains HVG data
    sce_iter <- sce_hvg
  }
  
  # calculate LISI scores
  lisi_scores <- list()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction and data extraction
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    y <- reducedDim(sce_iter, "PCA")
    cd <- colData(sce_iter)
    by <- c(
      LISI_c="group_id",
      LISI_g="cluster_id"
    )
    
    # LISI score calculation
    res <- sapply(by, \(.) {
        res <- compute_lisi(y, cd, .)
        n <- length(unique(cd[[.]]))
        (mean(res[[1]])-1)/(n-1)
    })
    
    lisi_scores[[selection]] <- data.frame(
      Selection = selection,
      LISI_g = unname(res["LISI_g"]),
      LISI_c = unname(res["LISI_c"])
    )
    
  }
  
  results <- do.call(rbind, lisi_scores)
  results$Selection <- factor(results$Selection, levels = results$Selection)
  return(results)
  
})
```

```{r}
p1 <- lisi_scores$patches_clu_con_full %>% 
  ggplot(aes(x = reorder(Selection, LISI_g), y = LISI_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_g", title = "Groups & Conditions (Group Mixing)")

p1
```

```{r}
p2 <- lisi_scores$patches_clu_con_full %>% 
  ggplot(aes(x = reorder(Selection, LISI_c), y = LISI_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_c", title = "Groups & Conditions (Condition Mixing)")

p2
```

```{r}
p3 <- lisi_scores$patches_con_full %>% 
  ggplot(aes(x = reorder(Selection, LISI_g), y = LISI_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_g", title = "Conditions (Group Mixing)")

p3
```

```{r}
p4 <- lisi_scores$patches_con_full %>% 
  ggplot(aes(x = reorder(Selection, LISI_c), y = LISI_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_c", title = "Conditions (Condition Mixing)")

p4
```

```{r}
# combine individual plots
lisi_plots <- list(p1, p3, p2, p4)

lisi_plots <- lapply(lisi_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(lisi_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "LISI Scores per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_lisi_full.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 10, dpi = 300)
combined_plot
```

```{r}
p1 <- lisi_scores$patches_clu_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, LISI_g), y = LISI_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_g", title = "Groups & Conditions (Group Mixing)")

p1
```

```{r}
p2 <- lisi_scores$patches_clu_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, LISI_c), y = LISI_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_c", title = "Groups & Conditions (Condition Mixing)")

p2
```

```{r}
p3 <- lisi_scores$patches_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, LISI_g), y = LISI_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_g", title = "Conditions (Group Mixing)")

p3
```

```{r}
p4 <- lisi_scores$patches_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, LISI_c), y = LISI_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "LISI_c", title = "Conditions (Condition Mixing)")

p4
```

```{r}
# combine individual plots
lisi_plots <- list(p1, p3, p2, p4)

lisi_plots <- lapply(lisi_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(lisi_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "LISI Scores per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_lisi_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 10, dpi = 300)
combined_plot
```

### PCR

```{r}
# define selections to be used
selections_clu_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "tMAV_sMAV",
  "tEN_sEN",
  "tMAD_sAD",
  "sMAV_tMAV",
  "sEN_tEN",
  "sAD_tMAD"
)

selections_con <- c(
  "all_genes",
  "random",
  "HVG",
  "tF_sPVE",
  "tF_sPBDS",
  "tPVE_sPVE",
  "sPVE_tF",
  "sPBDS_tF",
  "sPVE_tPVE",
  "common_sMAV",
  "common_sEN",
  "common_sAD",
  "sMAV_common",
  "sEN_common",
  "sAD_common"
)

all_sels <- unique(c(selections_clu_con, selections_con))
color_map <- setNames(qualitative_hcl(21, palette = "Dark 3"), all_sels)
```

```{r}
f <- function(x, id) {
  # regress PCs on 'id's
  y <- reducedDim(x, "PCA")
  ve <- attr(y, "varExplained")
  fit <- summary(lm(y~x[[id]]))
  # sum across coefficients of determination weighted
  # by variance explained & scale by total variance
  r2 <- vapply(fit, \(.) .$adj.r.squared, numeric(1))
  r2 <- r2[i <- r2 > 0]; ve <- ve[i]
  sum(ve*r2)/sum(ve)
}

pcr_scores <- lapply(all_selections, function(df) {
  
  # determine if df holds group and condition selections or only condition selections
  if ("tMAV" %in% colnames(df)) {
    # df contains group and condition selections
    selections <- selections_clu_con
  } else {
    # df contains only condition selections
    selections <- selections_con
  }
  
  # determine if df contains full or HVG data
  if (dim(df)[1] == dim(sce_full)[1]) {
    # df contains full data
    sce_iter <- sce_full
  } else {
    # df contains HVG data
    sce_iter <- sce_hvg
  }
  
  # calculate PCR scores
  pcr_scores <- list()
  for (selection in selections) {
    
    # set seed
    set.seed(42)
    
    # dimensionality reduction and data extraction
    sce_iter <- runPCA(sce_iter, subset_row=df[[selection]])
    ids <- c(
      PCR_g="cluster_id",
      PCR_c="group_id"
    )
    
    # PCR score calculation
    res <- sapply(names(ids), \(sta) f(sce_iter, ids[sta]))
    
    pcr_scores[[selection]] <- data.frame(
      Selection = selection,
      PCR_g = unname(res["PCR_g"]),
      PCR_c = unname(res["PCR_c"])
    )
    
  }
  
  results <- do.call(rbind, pcr_scores)
  results$Selection <- factor(results$Selection, levels = results$Selection)
  return(results)
  
})
```

```{r}
p1 <- pcr_scores$patches_clu_con_full %>% 
  ggplot(aes(x = reorder(Selection, PCR_g), y = PCR_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_g", title = "Groups & Conditions (Group Variance)")

p1
```

```{r}
p2 <- pcr_scores$patches_clu_con_full %>% 
  ggplot(aes(x = reorder(Selection, PCR_c), y = PCR_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_c", title = "Groups & Conditions (Condition Variance)")

p2
```

```{r}
p3 <- pcr_scores$patches_con_full %>% 
  ggplot(aes(x = reorder(Selection, PCR_g), y = PCR_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_g", title = "Conditions (Group Variance)")

p3
```

```{r}
p4 <- pcr_scores$patches_con_full %>% 
  ggplot(aes(x = reorder(Selection, PCR_c), y = PCR_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_c", title = "Conditions (Condition Variance)")

p4
```

```{r}
# combine individual plots
pcr_plots <- list(p1, p3, p2, p4)

pcr_plots <- lapply(pcr_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(pcr_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "PCR Scores per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_pcr_full.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 10, dpi = 300)
combined_plot
```

```{r}
p1 <- pcr_scores$patches_clu_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, PCR_g), y = PCR_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_g", title = "Groups & Conditions (Group Variance)")

p1
```

```{r}
p2 <- pcr_scores$patches_clu_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, PCR_c), y = PCR_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_c", title = "Groups & Conditions (Condition Variance)")

p2
```

```{r}
p3 <- pcr_scores$patches_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, PCR_g), y = PCR_g, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_g", title = "Conditions (Group Variance)")

p3
```

```{r}
p4 <- pcr_scores$patches_con_hvg %>% 
  ggplot(aes(x = reorder(Selection, PCR_c), y = PCR_c, fill = Selection)) +
    geom_col() +
    coord_flip() +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "PCR_c", title = "Conditions (Condition Variance)")

p4
```

```{r}
# combine individual plots
pcr_plots <- list(p1, p3, p2, p4)

pcr_plots <- lapply(pcr_plots, function(p) {
  p + theme(
    text = element_text(size = 16),
    plot.title = element_text(size = 18)
  )
})

combined_plot <- wrap_plots(pcr_plots, ncol = 2, guides = "collect") +
  plot_annotation(
    title = "PCR Scores per Selection",
    theme = theme(
      plot.title = element_text(
        hjust = 0.5,
        margin = margin(b = 30),
        size = 20,
        face = "bold"
      )
    )
  )

# save and show combined plot
ggsave("sel_pcr_hvg.png", combined_plot, path = "../../outs/sim/",
       width = 12, height = 10, dpi = 300)
combined_plot
```

### MI

```{r}
path <- "../../data/sim/05-mi/"
```

#### Embeddings

```{r}
emb <- read_csv(paste0(path, "embedding_mi.csv"), col_types = "cdddd")
head(emb)
```

```{r}
emb_experiments <- emb$experiment
color_map <- setNames(qualitative_hcl(5, palette = "Dark 3"), emb_experiments)
```

```{r}
p1 <- emb %>% 
  filter(experiment %in% c("emb_clu_con_hvg", "emb_con_hvg", "emb_con_hvg_dis")) %>% 
  ggplot(aes(x = reorder(experiment, mi_mean), y = mi_mean, fill = experiment)) +
    geom_col() +
    coord_flip() +
    geom_errorbar(aes(ymin = mi_mean - mi_std, ymax = mi_mean + mi_std), width = 0.2) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "MI", title = "Embeddings")

p1
```

#### Selections

```{r}
sel <- read_csv(paste0(path, "selection_mi.csv"), col_types = "ccdddd")
head(sel)
```

```{r}
sel_pairs <- unique(sel$selection_pair)
color_map <- setNames(qualitative_hcl(9, palette = "Dark 3"), sel_pairs)
```

```{r}
p2 <- sel %>% 
  filter(experiment == "sel_clu_con_full") %>% 
  ggplot(aes(x = reorder(selection_pair, mi_mean), y = mi_mean, fill = selection_pair)) +
    geom_col() +
    coord_flip() +
    geom_errorbar(aes(ymin = mi_mean - mi_std, ymax = mi_mean + mi_std), width = 0.2) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "MI", title = "Groups & Conditions")

p2
```

```{r}
p3 <- sel %>% 
  filter(experiment == "sel_con_full") %>% 
  ggplot(aes(x = reorder(selection_pair, mi_mean), y = mi_mean, fill = selection_pair)) +
    geom_col() +
    coord_flip() +
    geom_errorbar(aes(ymin = mi_mean - mi_std, ymax = mi_mean + mi_std), width = 0.2) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "MI", title = "Conditions")

p3
```

```{r}
p4 <- sel %>% 
  filter(experiment == "sel_clu_con_hvg") %>% 
  ggplot(aes(x = reorder(selection_pair, mi_mean), y = mi_mean, fill = selection_pair)) +
    geom_col() +
    coord_flip() +
    geom_errorbar(aes(ymin = mi_mean - mi_std, ymax = mi_mean + mi_std), width = 0.2) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "MI", title = "Groups & Conditions")

p4
```

```{r}
p5 <- sel %>% 
  filter(experiment == "sel_con_hvg") %>% 
  ggplot(aes(x = reorder(selection_pair, mi_mean), y = mi_mean, fill = selection_pair)) +
    geom_col() +
    coord_flip() +
    geom_errorbar(aes(ymin = mi_mean - mi_std, ymax = mi_mean + mi_std), width = 0.2) +
    theme_minimal() +
    theme(legend.position = "none") +
    scale_fill_manual(values = color_map) +
    labs(x = "", y = "MI", title = "Conditions")

p5
```









## Old Code

### Helper Functions

```{r}
# get top n genes for one score
get_top_n_genes <- function(df, score, n) {
  top_genes <- (df %>% 
    arrange(desc(.data[[score]])) %>% 
    slice_head(n = n))$gene_id
  
  return(top_genes)
}

# get top n genes for multiple scores
get_top_n_genes_scores <- function(df, scores, t_or_s, nt, ns) {
  results <- list()
  for (i in 1:length(scores)) {
    score <- scores[i]
    x <- t_or_s[i]
    if (x == "t") {
      results[[score]] <- get_top_n_genes(df, score, nt)
    } else if (x == "s") {
      results[[score]] <- get_top_n_genes(df, score, ns)
    }
  }
  return(results)
}

# get binary predictions for all scores in the selection list
compute_predictions <- function(selection_list, all_genes) {
  df <- data.frame(gene_id = all_genes)
  for (key in ls(selection_list)) {
    df[[key]] <- ifelse(all_genes %in% selection_list[[key]], 1, 0)
  }
  return(df)
}
```






























































